---
title: "Moving average"
author: "Steve Rolfe"
date: "26 October 2015"
output: html_document

Standard code to clear the system and install required libraries
```{r}
rm(list=ls())
library(ggplot2)
library(lubridate)
library(dplyr)
library(zoo)
```

Set up standard files names and the subset to plot
```{r}
setwd("C:/Users/Steve Rolfe/Google Drive/Phenomics/Water_Use")

input_file="./data/BR7_watering_data posix.csv"
plant_codes="./data/BR7_plant_codes.csv"

output_file="./output/BR7_water_out.csv"


#subset of data to plot using grep format
sset<-"BR7-0[1-5]1[1-3][1-4]"
```

We use a rolling average to smooth out the noise - this sets the window size for this average
To calculate the maximum value we use a larger window
Finally we set the tolerance of how close a value has to be to this maximum to be the maximum
```{R}
#rolling window size - for the rolling average of water data
wsz<-3
#maximum window size - for the number of points from which to calculate the maximum
msz<-6
#how close to the maximum value does a value have to be to be counted as a maximum
csz<-0.9
```


Set up standard headers for the files - these need to be edited to be consistent
The BR codes are suppled by IBERS but we should include the ASSYT codes where possible
Standard headers should be
BR_code,date_time, measurements (with informative names)
Custom code will be needed to translate the BR_codes into useful information

Read in the data and check that the required headers are there
```{r}
#read in the data
wdata<-read.csv(input_file,header=TRUE,stringsAsFactors = FALSE,row.names = NULL)
if(is.object(wdata)==FALSE) {stop(c("Can't open file:",input_file))}

#check that there the required headings are there and throw an error if not
if(is.na(match("BR_code",colnames(wdata)))){stop("BR_code missing from input file header")}
if(is.na(match("date_time",colnames(wdata)))){stop("date_time missing from input file header")}
head(wdata)

#this file contains info about the plants used
plant_codes<-read.csv(plant_codes,header = TRUE,stringsAsFactors = FALSE)
if(is.object(plant_codes)==FALSE) {stop(c("Can't open file:",plant_codes))}
```

We  deal with any specific coding here about names of plants, spores etc.
Use the file plant_codes to keep the information organised
```{r}
#dpi offset from first data point
#it's good to express everything as dpi but 1st data point might not be dpi
#set the date of infection
dpi_datetime<-"2015-7-6 00:00:00"

#now set the factors manually
wdata$BR_code<-as.factor(wdata$BR_code)

#and set the plants,spores,etc
#the plant code is the sixth character in the BR7 string
wdata$plant<-as.factor(substr(wdata$BR_code,6,6))
#the spore concentration code is the 8th character
wdata$spore<-as.factor(substr(wdata$BR_code,8,8))
#the repeat is the 9th
wdata$rep<-as.factor(substr(wdata$BR_code,9,9))
head(wdata)

#now merge the plant names with the data frame
wdata<-merge(wdata,plant_codes,by.x="plant",by.y="plant_code")
head(wdata)

#now do some standard time processing
#use difftime to get the days since inoculation
wdata$dpi<-as.numeric(difftime(wdata$date_time,dpi_datetime,units="days"))
#for a continuous x scale don't convert dpi into factors
head(wdata)
```

Now do the calculations on the selected subset of data
```{R}
subdata<-subset(wdata,grepl(sset,wdata$BR_code))

#use droplevels to get rid of 0 length levels from the subset - causes problems otherwise
subdata<-droplevels(subdata)
head(subdata)

#sort the data so that the rolling means and cumulative values are correct
subdata<-subdata[with(subdata,order(BR_code,dpi)),]
#get the rolling average by group (window = wsz)
subdata$ravg<-ave(subdata$diff_wt,subdata$BR_code,FUN=function(x) rollmean(x,k=wsz,fill = NA,na.pad=T))
#try the cumultative sum as well
subdata$cumsum<-ave(subdata$diff_wt,subdata$BR_code,FUN=function(x) cumsum(x))
```

Now do some plots
```{R}
#plot the rolling average and raw data - the facet wrap makes the plots for each plant separate
p1<-ggplot(data=subdata,aes(x=dpi,y=diff_wt,color=spore,group=BR_code))+geom_point()
p1<-p1+geom_line(data=subdata,aes(x=dpi,y=ravg,colour=spore,group=BR_code))
p1<-p1+scale_y_continuous(name="Weight g")
p1<-p1+facet_wrap(~plant_name)
p1

#just rolling average data
p2<-ggplot(data=subdata,aes(x=dpi,y=ravg,color=spore,group=BR_code))+geom_line()
p2<-p2+scale_y_continuous(name="Weight g")
p2<-p2+facet_wrap(~plant_name)
p2

#plot cumulative sums
p3<-ggplot(data=subdata,aes(x=dpi,y=cumsum,color=spore,group=BR_code))+geom_point()
p3<-p3+geom_line(data=subdata,aes(x=dpi,y=cumsum,colour=spore,group=BR_code))
p3<-p3+scale_y_continuous(name="Weight g")
p3<-p3+facet_wrap(~plant_name)
p3
```

Now get some parameters from the data
```{R}
#to find the peak values, then look in the rolling average data and get the nth maxima
#this creates a new sorted file (by group) with the NAs at the end
sort_subdata<-subdata[order(subdata$BR_code,subdata$ravg,na.last=TRUE,decreasing=TRUE),]

#create a vector to hold data by groups
grp<-group_by(sort_subdata,BR_code)
#get the mean of the msz th largest values
d<-summarise(grp,mean_ravg=mean(ravg[1:msz],na.rm=TRUE))
d
#merge
subdata<-merge(subdata,d,by="BR_code")

#now create a new column in subdata based on whether the ravg is less than the mean
subdata$lt_max<-subdata$ravg<(csz*d$mean_ravg[subdata$BR_code])
head(subdata)

#now get the first value which is less than csz*mean
pt1<-as.data.frame(aggregate(lt_max~BR_code,subdata,FUN=which.min))
#merge back into subdata
subdata<-merge(subdata,pt1,by="BR_code")
#the column names get odd so rename
subdata<-rename(subdata,lt_max_bool = lt_max.x)
subdata<-rename(subdata,lt_max_day = lt_max.y)
```

```{R}
#plot the time to maximum  as a boxplot
p1<-ggplot(subdata,aes(x=spore,lt_max_day))
p1<-p1+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p1<-p1+scale_y_continuous(name="Days to max weight")
p1<-p1+facet_wrap(~plant_name,ncol=5)
p1

#plot the maximum weight attained
p2<-ggplot(subdata,aes(x=spore,mean_ravg))
p2<-p2+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p2<-p2+scale_y_continuous(name="Max weight (g)")
p2<-p2+facet_wrap(~plant_name,ncol=5)
p2

write.csv(subdata,file=output_file)

```

